---
title: "Abalone Urchin IMTA"
author: "Aimee"
date: "04/04/2022"
output: html_document
keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("~/AbaloneIMTA")

```

This code is for the purpose of the data management module submission, but is beneficial for my honours project:
This code sorts through mortality weight data and plots a simple graph

## Setup
Make sure Abalone IMTA file is saved under documents in order for pathways to work

This chunk will read in each sheet

```{r load data, include=FALSE}
library(tidyverse)
library(readxl)
path<- "~/AbaloneIMTA/Data/abalone_measurements_raw.xlsx"
measurements_raw <- read_xlsx(path)
excel_sheets(path)
#[1] "Week 0" "Week 1" "Week 4" "Week 8"

week0raw<- read_excel(path, sheet = "Week 0")
week1raw<- read_excel(path, sheet = "Week 1")
week4raw<- read_excel(path, sheet = "Week 4")
week8raw<- read_excel(path, sheet = "Week 8")
```

## Tidy data
The shell length and width measurements were recorded using imaging software. 
The imaging software placed shell length in the first row and shell width in the second row for each individual.

This chunk of code will take the shell width measurements out and place them in a new column.
```{r tidy data, include=FALSE}
library(dplyr)
#subset data to isolate individual & length/width columns
week0<-week0raw[,c(2,3)]
week1<-week1raw[,c(2,3)]
week4<-week4raw[,c(2,3)]
week8<-week8raw[,c(2,3)]

names(week0)<-c("Individual","Length")
names(week1)<-c("Individual","Length")
names(week4)<-c("Individual","Length")
names(week8)<-c("Individual","Length")
#create dataset with only individual & width
week0Width<- week0 %>%
 filter(is.na(Individual))
week1Width<- week1 %>%
 filter(is.na(Individual))
week4Width<- week4 %>%
 filter(is.na(Individual))
week8Width<- week8 %>%
 filter(is.na(Individual))

#tidy width data set by removing individual column (since all values are NA)
#and rename column header to be just width
week0Width<-week0Width[,2]
names(week0Width) <- c("Width")
week1Width<-week1Width[,2]
names(week1Width) <- c("Width")
week4Width<-week4Width[,2]
names(week4Width) <- c("Width")
week8Width<-week8Width[,2]
names(week8Width) <- c("Width")
#remove rows with NA values from isolated data set i.e., deleting width measurements
week0<-week0 %>%
 filter(row_number() %% 2 == 1)
week1<-week1 %>%
 filter(row_number() %% 2 == 1)
week4<-week4 %>%
 filter(row_number() %% 2 == 1)
week8<-week8 %>%
 filter(row_number() %% 2 == 1)
#combine isolated data set with new width data set
week0<-bind_cols(week0,week0Width)
week1<-bind_cols(week1,week1Width)
week4<-bind_cols(week4,week4Width)
week8<-bind_cols(week8,week8Width)

#Add treatment column to weeks 1,4 & 8 to separate tank types
treatment_type1 <- rep("Treatment 1", length = 100)
treatment_type2 <- rep("Treatment 2", length = 100)
treatment_type3 <- rep("Treatment 3", length = 100)
treatment_type4 <- rep("Control 1", length = 100)
treatment_type5 <- rep("Control 2", length = 100)
treatment_type6 <- rep("Control 3", length = 100)
treatment <- c(treatment_type1,treatment_type2,treatment_type3,treatment_type4, treatment_type5, treatment_type6)
week1$Treatment <- treatment
week4$Treatment <- treatment
week8$Treatment <- treatment
week0$Treatment <- rep("Baseline", length = 300)

week0$Week <- rep("0", length=300)
week1$Week <- rep("1", length=600)
week4$Week <- rep("4", length=600)
week8$Week <- rep("8", length=600)


measurements_clean <- bind_rows(week0, week1, week4, week8)
```

Yay! data is tidy now!

## Save clean data
```{r, include = FALSE}
write.csv(week0,"~/AbaloneIMTA/Data/abalone_measurements_week0.csv")
write.csv(week1,"~/AbaloneIMTA/Data/abalone_measurements_week1.csv" )
write.csv(week4,"~/AbaloneIMTA/Data/abalone_measurements_week4.csv" )
write.csv(week8,"~/AbaloneIMTA/Data/abalone_measurements_week8.csv" )
write.csv(measurements_clean,"~/AbaloneIMTA/Data/abalone_measurements_clean.csv")
```


# Simple plot

Tank 0 is the baseline measurement measured at the start of the study (week 0)
Abalone shell length and width measurements taken at weeks 1,4 and 8 for the treatment and control tanks:
There are three treatment replicate tanks (Treatment 1, Treatment 2 & Treatment 3) and three control replicates (Control 1, Control 2, Control 3)

Treatment = abalone feed supplemented with sea urchin faeces
Control = abalone feed not supplemented with sea urchin faeces

Subset treatment & control data sets
```{r, include=FALSE}

treatment_data <- measurements_clean[measurements_clean$Treatment =="Treatment 1"|measurements_clean$Treatment =="Treatment 2"|measurements_clean$Treatment =="Treatment 3",]

control_data <- measurements_clean[measurements_clean$Treatment =="Control 1"|measurements_clean$Treatment =="Control 2"|measurements_clean$Treatment =="Control 3",]


```

```{r, echo=FALSE}
plot(treatment_data$Length, treatment_data$Width, xlab = "Shell length (cm)", 
     ylab = "Shell width (cm)", col = "blue")
points(control_data$Length, control_data$Width, col = "red")
legend("bottomright", c("Treatment", "Control"), col = c("blue", "red"), pch = c(21,21), bty = "n")
```
